\documentclass[a4paper]{article}

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[numbers]{natbib}

\title{How consistent are macroevolutionary and community ecology patterns of interspecific competition?}

\author{Marina C. Rillo} %, Mauro T. C. Sugawara, Rafael Lopes, \\ Michal Kucera, Thomas H. G. Ezard \& Brenno Cabella}

\date{\today}

\begin{document}
\maketitle

\begin{abstract}

A key ecological mechanism proposed to explain the lack of exponential diversification rates in the fossil record is competition among species. Under such a diversity dependent model, speciation rate declines with increasing standing species richness. However, if interspecific competition is an important driver of long-term evolution, we should expect competition to also structure modern communities. Two community-level patterns would support interspecific competition: (1) if species distributions result from competitive exclusion, we expect a pattern of allopatric species ranges. Alternatively, (2) if species coexist but still compete, we expect a negative correlation among species’ abundances through time. We used 35 population-level time-series encompassing 33 of the 47 extant species of the planktonic foraminifera clade. These time-series have mean length of 60 assemblage samples collected globally from sediment traps. We defined species pairs based on shell size and aspect ratio, and expected the most morphologically similar pairs to compete more strongly for resources than two randomly selected ones. Using co-occurrence analysis, we rejected the allopatric species ranges pattern since most species co-occur. Species’ abundances correlated more positively than negatively through time and no difference was found between morphologically similar species pairs and other pairs. Restricting analysis to the highest resolution time-series, we used Empirical Dynamical Models (EDM) that seeks to distinguish causality from correlation among species abundances time-series, considering seawater temperature time-series as well. The EDMs showed no causality between species time-series, and species continue to correlate positively even when considering temperature changes. In summary, we found no evidence of within-clade interspecific competition driving planktonic foraminifera population dynamics.  There is thus a mismatch between the processes inferred from shorter-term ecological patterns and those inferred from simplistic interpretations of macroevolutionary diversity dependence, especially when models fail to acknowledge that the changing biotic and abiotic environment alters the outcome of interspecific interactions.

\end{abstract}

%----------------------------------------------------------------------------------------

\section{Introduction}
\label{sec:introduction}
\input{time-introduction.tex}

%----------------------------------------------------------------------------------------

\section{Methods}
\label{sec:methods}
\input{time-methods.tex}

%----------------------------------------------------------------------------------------

\section{Results}
\label{sec:results}
\input{time-results.tex}

%----------------------------------------------------------------------------------------
\newpage

\section{Discussion}
\label{sec:discussion}
\input{time-discussion.tex}


%----------------------------------------------------------------------------------------

\label{Bibliography}
\bibliographystyle{abbrvnat} % Use the agsm, abbrvnat, unsrtnat BibTeX style for formatting the Bibliography
\bibliography{time-bibliography.bib} % The references (bibliography) information are stored in the file named "Bibliography.bib"

\end{document}  

